%$Header: /home/dashley/cvsrep/e3ft_gpl01/e3ft_gpl01/webprojs/pamc/gen_a/docs/manual/man_a/c_ehd0/c_ehd0.tex,v 1.2 2007/06/11 23:08:34 dashley Exp $

\chapter{The Employee Hours Database}

\label{cehd0}

\beginchapterquote{``TBD''}
                  {TBD}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%Section tag:  INT0
\label{cehd0:sint0}

Notes from discussions:

\begin{itemize}
\item Required (and minimum) granularity should be 0.25 hours.
\item Want to have the functionality of daily notes.  For each amount of time
      spend on each project for each day, employee may want to add notes (about
      what was worked on, what was accomplished, whatever).
\item Need the following reports:
      \begin{itemize}
      \item Date range.
      \item User(s).
      \item Project(s).
      \item Year-to-date (year-to-date totals are important for the user).
      \end{itemize}
      In the current system, can't do it by project or multiple users.
\item Should send e-mails when hours due and not complete.
\item Incomplete hours should show up in to-do list.  (Idea for this list:
      \emph{may} complete vs. \emph{should} complete vs. \emph{must}
      complete---perhaps one of those is optional.)
\item Locking:  The idea was floated that a user should have up to
      a week after time card submission to change data.  DTA response:
      bad idea---leads to possibility that data will be changed after it
      has moved upstream.  Proposed ideas:
      \begin{itemize}
      \item Time card makes a state change on user command.  It can't be
            used or contribute to reports until it has been officially submitted
            by the employee.  However, system should be able to record it until then.
      \item There is basic error checking that will block submission (or at least
            require a separate step) if the time card seems erroneous.
      \item Once a time card has been submitted/accepted, it will not be
            changeable any longer by the employee.  It will require administrator/supervisor
            approval to be unlocked again.
      \end{itemize} 
\item No access by one employee to another's data, except:
      \begin{itemize}
      \item Administrators.
      \item Those above the employee in the chain of command.  (Note:  this implies that
            the application know the chain of command.  Chain of command should also
            probably include weak reporting or matrix reporting relationships.  Cycles
            should be prevented.)
      \end{itemize}
\item YV has details about how project numbers are assigned.
\item Unclear whether daily submissions or weekly or both should be supported, and how.
      Some employees prefer daily, some weekly.  Will need to cache hours until weekly,
      perhaps.  Need to think further about supporting users who like day-by-day vs.
      weekly.      
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\begin{figure}[!b]
\noindent\rule[-0.25in]{\textwidth}{1pt}
\begin{tiny}
\begin{verbatim}
$RCSfile: c_ehd0.tex,v $
$Source: /home/dashley/cvsrep/e3ft_gpl01/e3ft_gpl01/webprojs/pamc/gen_a/docs/manual/man_a/c_ehd0/c_ehd0.tex,v $
$Revision: 1.2 $
$Author: dashley $
$Date: 2007/06/11 23:08:34 $
\end{verbatim}
\end{tiny}
\noindent\rule[0.25in]{\textwidth}{1pt}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%$Log: c_ehd0.tex,v $
%Revision 1.2  2007/06/11 23:08:34  dashley
%Edits.
%
%Revision 1.1  2007/06/11 22:40:21  dashley
%Initial checkin.
%
%End of $RCSfile: c_ehd0.tex,v $.
