\chapter{Acknowledgements}

I am very grateful for all
of the assistance I have received with this set of works (the book,
the tool set, and supplemental materials).  I would
like to extend my gratitude to the following individuals
and institutions.

For assistance in procuring \LaTeXe{} (\emph{4allTex} Version 5), I am
grateful to Mr. Erik Frambach, Mr. Gerard van Nes, and the Dutch
\TeX{} User's Group.

For assistance in learning how to use \LaTeX{} and \TeX{}, I am
especially grateful to Mr. Tobias Oetiker\footnote{Author of
\emph{The Not So Short Introduction To \LaTeXe{}}, which is
available on the Web.},
who provided detailed e-mail answers to all of
my questions.  Mr. Oetiker also provided the source code for his
document on the Web, and I was able to use this as a template to
jump-start me in the direction of an aesthetically pleasing
\LaTeX{} book.  I am also grateful to the Dutch \TeX{}
User's Group, the UK \TeX{}
User's Group (UKTUG), and many helpful posters on the 
\texttt{comp.text.tex} newsgroup
\cite{bibref:n:comptexttex} for support.

For hosting this project (the book, the tool set, and all supporting
materials), I am very grateful to \index{SourceForge}\emph{SourceForge}
(\texttt{http://www.sourceforge.net})---coordinating and 
distributing this project simply would not have been possible without
their support.  At SourceForge, I am very grateful to Mr.
Jacob Moorman for technical support and assistance on many,
many occasions.

I would like to acknowledge the contribution of \index{GNU}GNU 
(\texttt{http://www.gnu.org}), 
\index{Free Software Foundation (FSF)}The Free Software Foundation 
(\texttt{http://www.fsf.org}), and all of the open-source
developers in the world who have produced many excellent and
helpful products, including
\LaTeX{} (used to write this book) and 
\index{Tcl}\emph{Tcl} (the basis
for much of \emph{The ESRG Tool Set}).

Finally, I would like to gratefully acknowledge the
topic-specific contributions of the individuals who are
acknowledged at the end of each chapter.

%End of ucbka_acknowledgments.tex

