%$Header: /home/dashley/cvsrep/e3ft_gpl01/e3ft_gpl01/webprojs/pamc/gen_a/docs/manual/man_a/comps/workprfa.tex,v 1.9 2009/10/31 19:02:03 dashley Exp $
\chapter{Preface}

The product described in this document, 
\emph{\productbasename{} Version \productversion{}}
(also called \emph{\productbasename{}-\productversion{}}),
is a free open-source server-based collaboration solution for 
product development.  

\emph{\productbasename{}-\productversion{}}
is a recipe for transforming 
a \emph{*nix} server into a 
product development collaboration platform.
\emph{\productbasename{}-\productversion{}} does provide
original content, but a large part of the functionality of
\emph{\productbasename{}-\productversion{}} is built
on other open-source products such as 
\emph{*nix}, \emph{cvs}, \emph{svn}, \emph{PHP}, 
\emph{MySQL}, and \emph{putty}.

\emph{\productbasename{}-\productversion{}} provides the
following functionality:

\begin{itemize}
\item Secure access to \emph{*nix} shell accounts and
      SFTP, secured by \emph{SSH} and \emph{CryptoCard} tokens.
\item Personal web space, secured by \emph{CryptoCard} tokens
      used in conjunction with an \emph{Apache} authentication
      module.
\item Version control, implemented via 
      \emph{cvs} and \emph{svn} and secured
      via secure port forwarding and
      \emph{putty}.
\item A general web database platform.

   \begin{itemize}
   \item The platform is secured with \emph{CryptoCard} tokens,
         static passwords, IP restrictions, and \emph{https}.
   \item The platform as installed provides substantial
         useful functionality, including:

         \begin{itemize}
         \item User authentication and management.
         \item E-mail notifications.
         \item Employee time tracking and hours reporting.
         \item Viewing of version control archives.
         \item Document management.
         \item Software defect tracking.
         \item Automatic database maintenance.
         \end{itemize}

   \item The platform is designed to be easily extended by allowing
         the addition of custom applications that operate in the same
         authentication framework.
   \end{itemize}
\end{itemize}

\emph{\productbasename{}-\productversion{}} can be used as-is, or individual
components can be harvested for use in other applications.  For example,
the compiled \emph{C} program and \emph{PHP} functions that allow \emph{CryptoCard} tokens
to be used from \emph{PHP} may be useful in other applications.

\emph{\textbf{Part I: General Information}} provides introductory
and general information about the product.

\emph{\textbf{Part II: Technical Information}} provides a detailed
technical description of the product and its theory of operation.

\emph{\textbf{Part III: Installation and Maintenance}} provides an installation
guide for the product and instructions for maintaining the product.  
\emph{\productbasename{}-\productversion{}}
is designed for use only on \emph{*nix} platforms.

\emph{\textbf{Part IV: User Information}} provides a 
non-technical user manual for the product.

\emph{\textbf{Part V: Procedures and Checklists}} provides a 
set of procedures and checklists to accomplish tasks where the
procedure or checklist didn't fit well into any other part of this
document.

\emph{\textbf{Part VI: Appendices, Bibliography, and Index}} provides
glossaries, references, and an index.
Individuals, products, companies, websites, and Internet newsgroups
are cited in the same framework
as traditional references in order to provide the reader with more
resources to obtain information.

The \emph{\productbasename{}-\productversion{}} software and all
associated materials are provided under open-source licenses; so all source
code is available and the product can be modified and/or distributed with
minimal restrictions.

Please feel free to contact me at \texttt{dashley@gmail.com} with
any suggestions for the \emph{\productbasename{}-\productversion{}}
product or associated documentation.  If you require
assistance with the product, please refer to
\S{}\ref{ciov0:soas0} (p. \pageref{ciov0:soas0}).

I hope that this product is useful to you and that you
find it easy to install, use, maintain, and modify.

\vspace*{0.5in}

\noindent{}\hspace*{75mm}David T. Ashley                \\
           \hspace*{75mm}Marshall, Michigan, USA        \\
           \hspace*{75mm}November, 2009                 \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent\begin{figure}[!b]
\noindent\rule[-0.25in]{\textwidth}{1pt}
\begin{tiny}
\begin{verbatim}
$RCSfile: workprfa.tex,v $
$Source: /home/dashley/cvsrep/e3ft_gpl01/e3ft_gpl01/webprojs/pamc/gen_a/docs/manual/man_a/comps/workprfa.tex,v $
$Revision: 1.9 $
$Author: dashley $
$Date: 2009/10/31 19:02:03 $
\end{verbatim}
\end{tiny}
\noindent\rule[0.25in]{\textwidth}{1pt}
\end{figure}

%$Log: workprfa.tex,v $
%Revision 1.9  2009/10/31 19:02:03  dashley
%Edits.
%
%Revision 1.8  2009/10/29 17:46:12  dashley
%Edits.
%
%Revision 1.7  2009/10/29 04:52:52  dashley
%Edits.
%
%Revision 1.6  2007/07/11 01:36:52  dashley
%Edits.
%
%Revision 1.5  2007/06/06 00:32:07  dashley
%Edits.
%
%Revision 1.4  2007/06/05 15:47:12  dashley
%Structural edits.
%
%Revision 1.3  2007/06/04 06:26:36  dashley
%Edits.
%
%Revision 1.2  2007/06/03 04:57:47  dashley
%Checkin to be sure that changes wrapped up in weekly TAR.GZ file.
%
%Revision 1.1  2007/04/10 23:44:47  dashley
%Initial checkin.
%
%End of $RCSfile: workprfa.tex,v $.
