%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{The \emph{Clike} Programming Language}
\label{ccik2}

The \index{Clike@\emph{Clike}}\emph{Clike}\footnote{Mnemonic: \emph{C}-\emph{l}ike.}
programming language is a ``glue'' language to integrate tools.  This chapter describes
the language in detail.

\emph{Clike} is designed to be similar enough to
\index{C programming language@\emph{C} (programming language)}\emph{C}
(with some \index{C++ programming language@\emph{C++} (programming language)}\emph{C++}
ideas included) that any \emph{C}/\emph{C++} programmer can understand and
use the language immediately.

Entire \emph{Clike} programs contained in source files can be run; but individual
statements can also be executed interactively.  This makes a tool aggregation that
includes the \emph{Clike} intepreter often useful as a general utility.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rationale for \emph{Clike}}
\label{ccik2:swno0}

There are many capable scripting languages available, including
\index{Python (scripting language)@\emph{Python} (scripting language)}\emph{Python},
\index{Perl (scripting language)@\emph{Perl} (scripting language)}\emph{Perl}, and
\index{Tcl (scripting language)@\emph{Tcl} (scripting language)}\emph{Tcl}.
Why was \index{Clike (scripting language)@\emph{Clike} (scripting language)}\emph{Clike} developed?

The primary use of \emph{Clike} is twofold:

\begin{itemize}
   \item Creating and manipulating source code and data directly used in the construction
         of embedded systems.
   \item Embedded systems engineering and development.
\end{itemize}

The first use, in particular, demands high reliability.  A software defect
in \emph{Clike} or the underlying tools could lead to a software defect
in an embedded system.

To varying degrees, 
\index{Python (scripting language)@\emph{Python} (scripting language)}\emph{Python},
\index{Perl (scripting language)@\emph{Perl} (scripting language)}\emph{Perl}, and
\index{Tcl (scripting language)@\emph{Tcl} (scripting language)}\emph{Tcl} have the 
following disadvantages which impact reliability:

\begin{itemize}
   \item Dissimilarity to \emph{C}.  (Additional learning effort.
         \emph{C} is the dominant language for embedded systems
         development, and virtually every software developer is fluent.)
   \item Inability to statically link.  (In very conservative build processes,
         it is common to calculate the cryptographic hashes of executables and
         compare them to the expected values to ensure that the expected tools are
         being used in the build.  Software with dynamic library dependencies
         renders this process less reliable.)
   \item Difficulty debugging.  (\emph{Clike} has several features to facilitate
         debugging, including logging and the built-in ability to emit bytecode and
         system state.)
   \item Inability to unit test software components.  (The \emph{Clike} interpreter
         is built to be unit-tested.)
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General Features of the Language and the Interpreter}
\label{ccik2:sgft0}

\emph{Clike} has the following features and characteristics.

\begin{itemize}
   \item Minimalist nature.
         \begin{itemize}
            \item The language is intended for basic ``glue'' scripting, and so is Turing-complete
                  but doesn't have many common features of high-level languages.
                  Missing features\footnote{Some of these features may be added later.}
                  include:
                  \begin{itemize}
                     \item Support for non-ASCII character sets in \emph{Clike} source code.
                     \item Direct support for processing Unicode in files.
                     \item Object-oriented features.
                     \item Constant variables.
                     \item The \emph{switch()}\footnote{\emph{switch()} must be coded using
                           \emph{else-if}.} statement.
                     \item Programmatic interpretation of \emph{Clike} scripts.\footnote{All source code
                           must be available at the start of the program.  The source code is byte-coded
                           once at the start of the program, and no elements may be byte-coded later.}
                  \end{itemize}
         \end{itemize}
   \item Similarity to
         \index{C (programming language)@\emph{\emph{C} (programming language)}}\emph{C}.
         \begin{itemize}
            \item The similarity (plus the minimalist nature of the language) should allow
                  a programmer to learn the language quickly.
         \end{itemize}
   \item Support for local multi-threading.
      \begin{itemize}
         \item A significant use of the \emph{Clike} language will be for scripting embedded
               builds.
               \begin{itemize}
                  \item Such builds for large embedded systems now typically require over 10 minutes
                        on a personal computer.
                  \item Using all cores available on a personal computer is necessary to minimize
                        build time.
               \end{itemize}
      \end{itemize}
   \item Sub-optimal performance.
      \begin{itemize}
         \item The \emph{Clike} interpreter is not designed to be fast (unlike \emph{Python} or
               certain other scripting languages).
               \begin{itemize}
                  \item However, it is anticipated that the highest-frequency operations will occur
                        in compiled code, either within the \emph{Clike} interpreter or in spawned
                        programs, so the performance of the \emph{Clike} interpreter itself
                        won't have much of an effect on overall execution times for typical
                        use cases.
               \end{itemize}
      \end{itemize}
   \item Designed for reliability and debugging.
      \begin{itemize}
         \item Most components of the tool set are unit-tested with MC/DC coverage.
         \item The \emph{Clike} interpreter is designed to be simple, at the possible
               expense of speed.
         \item The \emph{Clike} interpreter and any aggregated tools are statically linked, and so the
               cryptographic hash of the resulting executable can be used to ensure that the intended
               version of the executable is being used.
         \item There are adequate debugging options, including:
               \begin{itemize}
                  \item Launching the interpreter with a high verbosity and/or logging level so that
                        issues can be found.
                  \item The availability of commands the output the internal state, including variables
                        and call stacks.
                  \item The ability to debug under development tools.
               \end{itemize}
      \end{itemize}
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\emph{Clike} Script and Statement Execution}
\label{ccik2:ssfi0}

A \emph{Clike} script or statement is always executed in one of the following three ways:

\begin{itemize}
   \item A tool aggregation containing a \emph{Clike} interpreter is run with
         the name of the file containing a script\footnote{This is the master file---it may
         include other files as described in \S{}\ref{ccik2:ssff0}.} specified as a command-line
         parameter.  In this case, the tool aggregation program ends when the script
         ends.
   \item A single \emph{Clike} statement---always an assignment or expression evaluation
         possibly involving function calls---is run interactively.
   \item The \emph{source()} function is invoked interactively, and the 
         \emph{Clike} interpreter executes the master file specified as the argument
         specifed to the \emph{source()} function.  In this case, no information
         is transferred from the interactive environment to the executed script, and the
         tool aggregation program ends when the script ends.
\end{itemize}

There is no provision for allowing one script to run another as part of the same
interpreter invocation, for allowing source code to be constructed and compiled to
bytecode during the execution of a \emph{Clike} script, or for allowing bytecode
to be modified once a script's bytecode has begin execution.

File inclusion is supported (\S{}\ref{ccik2:ssff0}), but this does not change
the paradigm.  A single
``master'' script, which possibly includes files, is compiled to bytecode.
The bytecode is thereafter never modified during that invocation of the 
aggregation containing the \emph{Clike} interpreter.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{File Inclusion}
\label{ccik2:ssff0}

\emph{Clike} includes the \emph{include()} and \emph{includeonce()} directives.

\emph{includeonce()} is identical to \emph{include()} except that it automatically
prevents multiple inclusion.  (Restrictions:  same string, exactly.)

These directives are evaluated before byte-coding, so the arguments cannot be calculated
values.

These may be placed anywhere within a file on their own line. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Comments}
\label{ccik2:scom0}

Comments may be done in either the \emph{C} or \emph{C++} styles.

\emph{C} comments may be nested, with the restriction of matching comment delimiters.  (This
decision was made due to the lack of preprocessing ability.)

\emph{C++} comments take precedence over \emph{C} comments in that a \emph{C++} comment
may block a \emph{C} comment, but not vice-versa (see Figs. \ref{fig:ccik2:scom0:01}
and \ref{fig:ccik2:scom0:02}).

\begin{figure}
\begin{verbatim}
/*
//  /* The C-style comment delimiter to the left is ignored,
//  /* and the comment block here will not cause errors.
*/
\end{verbatim}
\caption{\emph{C++}-Style Comment Preventing Recognition of a \emph{C}-Style Comment}
\label{fig:ccik2:scom0:01}
\end{figure}

\begin{figure}
\begin{verbatim}
//The right C-style delimiter below will not be recognized, and
//the code will likely cause a parsing error.
/* //C-style delimiter to the right not seen. */
\end{verbatim}
\caption{\emph{C}-Style Comment \emph{Not} Preventing Recognition of a \emph{C++}-Style Comment}
\label{fig:ccik2:scom0:02}
\end{figure}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Language Version Number Declaration}
\label{ccik2:slvn0}

The \emph{clikeversion()} directive must be the first parsable item in the master
source file, and must provide a version number that matches the interpreter.

This directive is provided and required because the language is in a state of flux.
At this time, it is necessary to strongly pair scripts with interpreter versions.

For now, the version is \emph{0.1a}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keywords and Lexical Tokens}
\label{ccik2:skwd0}

The keywords and lexical tokens:

\begin{itemize}
   \item Expression order of evaluation
      \begin{itemize}
          \item \emph{(}
          \item \emph{)}
      \end{itemize}
   \item Statement blocks
      \begin{itemize}
          \item \emph{\{}
          \item \emph{\}}
      \end{itemize}
   \item Assignment
      \begin{itemize}
          \item \emph{=}
      \end{itemize}
   \item Comparison
      \begin{itemize}
          \item \emph{==}
          \item \emph{!=}
          \item \emph{\textless}
          \item \emph{\textless{}=}
          \item \emph{\textgreater}
          \item \emph{\textgreater{}=}
      \end{itemize}
   \item Function definition
      \begin{itemize}
          \item \emph{proc}
          \item \emph{return}
      \end{itemize}
   \item Looping
      \begin{itemize}
          \item \emph{for}
          \item \emph{while}
          \item \emph{break}
          \item \emph{continue}
      \end{itemize}
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variables}
\label{ccik2:svar0}

TBD.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Literals}
\label{ccik2:slit0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Primitive Data Types}
\label{ccik2:ssdt0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Explicit Type Conversion}
\label{ccik2:stcv0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Operators}
\label{ccik2:sops0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Arrays and Structure Field Selection}
\label{ccik2:sasf0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statements}
\label{ccik2:ssta0}

Need to include the semicolon and comma operators.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Blocks}
\label{ccik2:sblk0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Logical Expressions}
\label{ccik2:slex0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The \emph{if()} Statement}
\label{ccik2:sifs0}

Need to include else and elseif here.

No switch() statement.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The \emph{while()} and \emph{for()} Loops}
\label{ccik2:swfl0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The \emph{do-while()} Loop}
\label{ccik2:sdwl0}

TBD.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functions}
\label{ccik2:sfun0}

Need to include pass-by-value and pass-by-reference.

For pass by value, value won't have a copy made unless there is an attempt to modify the
variable.  This is implicit const.

Need to include recursion.

Need to include variable-length parameter lists.

Need to include default parameters.

Every function returns a value.  If no return statement occurs, 0 is returned automatically.

Need to have the \emph{global} keyword to allow access to anything not in the function.

No notion of static.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Errors and Error Termination}
\label{ccik2:seet0}

Decided that will have the \emph{try\{\}} and \emph{catch()\{\}}
syntax like C++.

Multiple \emph{catch} clauses my be sequential.  Search performed in order.  Must immediately
follow \emph{try}.

Need to have a list of error types, codes, etc.

Need to have a \emph{throw}.

Errors percolate up the calling tree to first applicable \emph{catch}.  Anything uncaught terminates program.

Out of memory is fatal.

Mutexes and file handles (as well as, naturally, variables) may be left in an undesired state if
not properly handled.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Built-In Functions}
\label{ccik2:sbif0}

Every built-in function should begin with an underscore.  No user-defined function may begin
with an underscore.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Built-In Constants}
\label{ccik2:sbic0}

\_PFORM.
\_CMDLINEPARS.
\_ENV.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Internal Representations}
\label{ccik2:sirp0}

This should possibly go in a design section.

Need to have the following data structures:

\begin{itemize}
\item Byte-coded source deck.\\
      Fully formed by the time threads might start.\\
      Read-only.
\item Global variables.\\
      Shared among all threads.\\
      The classic IPC problems may exist, but not in a catastrophic form (the low-level operations,
      such allocation, assignment, moving pointers, etc. are protected from threading issues).  But,
      above that level, RMW and race conditions may exist.
\item Call stack.\\
      Per thread.\\
      Contains function calls, return addresses, parameters, local variables, and return values that were
      pushed onto the stack or exist on the stack.\\
      Some counterintuitive protection may be necessary with pass-by-reference.
\item Global constants.\\
      Shared among all threads.
      Fully formed by the time threads might start.
      Read-only.
\end{itemize}

Need to have a way to dump the internal structures for debugging.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-Threading (Concurrency)}
\label{ccik2:scnc0}

Threads, mutexes, and condition variables may be created only once; and must be created all at the
same time.

All threads must begin at UserThreadFunc() (as opposed to the primary thread, which begins at
\emph{main()}.

Effectively, the worker thread pool approach is enforced.

Mutexes and so on need to be tracked so that the state can be determined on debug dumps.

Each thread has access to the same pool of global variables (no notion of thread-local storage).

Mutexes should be used for protection.

Automatic mechanisms prevent the most egregious violations that would result in memory leaks, allocation problems, etc.

Certain functions will be inherently atomic (such as \emph{puts}).

There should be only three total mutexes:
\begin{itemize}
\item The mutex used to keep certain critical data structure operations atomic (not available to
      \emph{Clike} programs).
\item The mutex used to keep certain function calls atomic (not available to \emph{Clike} programs).
\item The user mutex, used to protect critical sections between threads (manipulated by Clike programs).
\end{itemize}

The order of obtaining the mutexes will always be inherently from the last element of the enumeration
above to the first, so deadlock should not be possible without a programming mistake.

The main thread should process any sort of a break signal, then set an internal variable so that
the worker threads can come back to a known state.

Unsure of how to protect for deadlock on the user mutex.  For initial versions may allow a hang,
then for later versions fix it.  Perhaps can put a counter in the break handler where if there are
too many break events it gets more forcible.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Command Prompt Semantics}
\label{ccik2:scps0}

When typing a command at the command prompt, a semicolon is not required, but a continuation character can be
used.

Need to establish the correspondence between tools and the Clike namespace.  Need to do better with errors, because otherwise the
command-line would be burdened with those.  Will need a \emph{catch} it seems.

Anything that would have been caught would be a printed error.

The parenthesis on a function can be omitted.

\emph{tl.tool.subcommand(pars)}
\emph{tl.tool.subcommand pars}

The scripted form of a command invocation returns the outputs as an associative array.  In the
case of command-line invocation, these will be displayed automatically.  In the case of scripted
invocation, these will be assigned to a variable.

If there are errors, these will be displayed.  Essentially, this would be an uncaught
exception.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

