%$Header: /cvsroot/esrg/sfesrg/esrgpcpj/hyreach/doc/hyreachm/s_glo0/s_glo0.tex,v 1.4 2002/01/29 17:04:00 dtashley Exp $
%
\section{Glossary Of Terms}
\label{sglo0}

\begin{glossaryenum}
\item \textbf{CHM}\index{CHM} \\
      \emph{C}omposite \emph{h}ybrid \emph{m}achine.  A timed automaton which is
      created by \swname{} by combining the state-spaces of all EHMs supplied as
      input using a well-defined mathematical operation.  In general, each location
      of a CHM corresponds to one location from each EHM.
\item \textbf{configuration}\index{configuration} \\
      See \emph{location}.
\item \textbf{EHM}\index{EHM} \\
      \emph{E}lementary \emph{h}ybrid \emph{m}achine.  A timed automaton supplied
      as input to \swname{}.
\item \textbf{location}\index{location} \\
      The discrete component of an EHM's state.  The state of an EHM is divided into
      two components:
      \begin{itemize}
      \item The discrete state, which is usually shown on a state transition diagram
            as a circle or a rectangle.
      \item The values of clocks which the EHM sets and tests.  These values are
            conceptually continuous, but for machine representation reasons are
            confined to integers.
      \end{itemize}
      The \emph{location} is the disctete state only.
\item \textbf{on-the-fly}\index{on-the-fly} \\
      Creating (dynamically allocating) the locations of a CHM as needed during 
      verification, rather than in
      advance.  For analyzing reachability, creating locations of a CHM before verification
      begins has performance disadvantages for three reasons:
      \begin{itemize}
      \item A large number of CHM locations may not be reachable, and so are not used
            during the verification.  Creating these locations is a wasted motion.
      \item In the event that an illegal location is reachable, in most cases
            creating all CHM locations
            in advance will greatly slow down finding 
            that the illegal location is reachable.  Because a model with illegal
            states reachable usually requires correction, this will slow down the 
            development cycle in which \swname{} is being used.
      \item For models with a very large state space, not all symbolic information for 
            all locations can be 
            kept in memory at one time (and it is not necessary to do so).  For large
            models, it is not
            possible to create all locations in advance.
      \end{itemize}
\end{glossaryenum}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\begin{figure}[!b]
\noindent\rule[-0.25in]{\textwidth}{1pt}
\begin{tiny}
\begin{verbatim}
$RCSfile: s_glo0.tex,v $
$Source: /cvsroot/esrg/sfesrg/esrgpcpj/hyreach/doc/hyreachm/s_glo0/s_glo0.tex,v $
$Revision: 1.4 $
$Author: dtashley $
$Date: 2002/01/29 17:04:00 $
\end{verbatim}
\end{tiny}
\noindent\rule[0.25in]{\textwidth}{1pt}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%$Log: s_glo0.tex,v $
%Revision 1.4  2002/01/29 17:04:00  dtashley
%Version control info added, and minor edits.
%
%Revision 1.3  2001/10/31 04:31:37  dtashley
%Nightly safety checkin.
%
%Revision 1.2  2001/10/01 06:08:09  dtashley
%Basic hacks in place for a workable document.
%
%Revision 1.1  2001/09/26 02:31:14  dtashley
%Initial checkin, and some edits of main TEX file.
%
%End of S_GLO0.TEX.
