%$Header: /cvsroot/esrg/sfesrg/esrgpcpj/doc/engman01/c_prf0/c_prf0.tex,v 1.4 2002/06/26 01:01:49 dtashley Exp $
\chapter{Preface}

\emph{The \tsname{} Version \tsversion{}} is a research-oriented tool set
developed to be useful in microcontroller and embedded control work.
The tool set contains a companion research work, the book entitled
\emph{\compbooktitlelong}.

Although the companion work is shamelessly and
deliberately somewhat of a catchall, it did not
seem useful to include tool set build instructions and other tool
engineering details in that work.  For the most part, building
\emph{The \tsname{}} from source code or modifying or contributing to
it is something that most users and casual readers will not
desire to do.  For this reason, those details are not
included in the companion work; instead, tool set engineering 
details here in this engineering manual.

This engineering manual is intended for individuals who desire
to:

\begin{itemize}
\item Build \emph{The \tsname{}} from source code.
\item Modify the tool set.
\item Collaborate with the ESRG to extend the tool set. 
\end{itemize}

Please direct all comments and suggestions about this engineering
manual to \texttt{toolbugs@esrg.org}.

\vspace{4mm}

\noindent{}---Dave Ashley (\texttt{dtashley@esrg.org}) \\
06/25/02, Detroit, Michigan, USA.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\begin{figure}[!b]
\noindent\rule[-0.25in]{\textwidth}{1pt}
\begin{tiny}
\begin{verbatim}
$RCSfile: c_prf0.tex,v $
$Source: /cvsroot/esrg/sfesrg/esrgpcpj/doc/engman01/c_prf0/c_prf0.tex,v $
$Revision: 1.4 $
$Author: dtashley $
$Date: 2002/06/26 01:01:49 $
\end{verbatim}
\end{tiny}
\noindent\rule[0.25in]{\textwidth}{1pt}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% $Log: c_prf0.tex,v $
% Revision 1.4  2002/06/26 01:01:49  dtashley
% Edits, safety checkin.
%
% Revision 1.3  2002/06/25 02:44:16  dtashley
% Edits.
%
% Revision 1.2  2002/06/24 18:09:19  dtashley
% Formatting of log corrected.
%
% Revision 1.1  2002/06/24 18:08:44  dtashley
% Initial checkin.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%End of file C_PRF0.TEX.
